---
title: "Data Analysis of T-Cell Dysfunction in the Glioblastoma Microenvironment is Mediated by Myeloid Cells Releasing Interleukin-10"
author: "Banupriya Jayarama Reddy"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Load required library

```{r}
#===============================================================================
# Library
#===============================================================================
library(tidyverse)
library(treemapify)

```


Import the data into objects but somehow the factor was not kept from the `data-preprocessor.R` script so have to tidy up the data again for further data analysis.


```{r}
#===============================================================================
# Import
#===============================================================================
# load the data into individual objects
tumor_data <- read.csv(file = "tumor_data.csv")
t_cell_data <- read.csv(file = "t_cell_data.csv")
# load the first six line of data just to double check
head(tumor_data)
head(t_cell_data)

#===============================================================================
# Tidy
#===============================================================================
# clean up the data by making them into factor, somehow factor in data-preprocessor.R did not work
# tumor data
tumor_data$Cell_ID_Myeloid <- as.factor(tumor_data$Cell_ID_Myeloid)
tumor_data$Cell_ID_Tumor <- as.factor(tumor_data$Cell_ID_Tumor)
tumor_data$Cell_type <- as.factor(tumor_data$Cell_type)
# use tibble to tidy up the data
tumor_data <- as_tibble(tumor_data)

# t cell data
t_cell_data$Cell_ID_Myeloid <- as.factor(t_cell_data$Cell_ID_Myeloid)
t_cell_data$Cell_ID_T_Cell <- as.factor(t_cell_data$Cell_ID_T_Cell)
t_cell_data$Cell_type <- as.factor(t_cell_data$Cell_type)
# use tibble to tidy up the data
t_cell_data <- as_tibble(t_cell_data)

# generate summary and correlation to give an overview
summary(t_cell_data)
summary(tumor_data)
```

Reproducing the density graph like in the Figure 6c for the purpose of validating the data.

```{r}
#===============================================================================
# Validate
#===============================================================================
# tumor
# compute density
density_tumor_distance_hpositive <- density(tumor_data$Distance[tumor_data$Cell_type == "H+"])
density_tumor_distance_hnegative <- density(tumor_data$Distance[tumor_data$Cell_type == "H-"])
ggplot(tumor_data, aes(x = Distance, color = Cell_type, fill = Cell_type)) +
  geom_density(alpha = 0.5) +
  labs(title = "Distance to Tumor cells", y = "Density")
#  xlim(c(min(density_tumor_distance_hpositive$x, density_tumor_distance_hnegative$x),
#         c(max(density_tumor_distance_hpositive$x, density_tumor_distance_hnegative$x)))) +
#  scale_fill_discrete(name = "Cell type", labels = c("H+", "H-"))
#dev.off()

# t cell
# compute density
density_t_cell_distance_hpositive <- density(t_cell_data$Distance[t_cell_data$Cell_type == "H+"])
density_t_cell_distance_hnegative <- density(t_cell_data$Distance[t_cell_data$Cell_type == "H-"])
# plot density graph
#png(file = "./density_ggplot_distance_to_t_cells.png")
# normal plot
#plot(density_t_cell_distance_hpositive, lwd = 2, col = "blue", main = "Distance to T cells")
#polygon(density_t_cell_distance_hpositive, col = rgb(0, 0, 1, alpha = 0.5))
#lines(density_t_cell_distance_hnegative, lwd = 2, col = "red")
#polygon(density_t_cell_distance_hnegative, col = rgb(1, 0, 0, alpha = 0.5))
# ggplot
ggplot(t_cell_data, aes(x = Distance, color = Cell_type, fill = Cell_type)) +
  geom_density(alpha = 0.5) +
  labs(title = "Distance to T-cells", y = "Density")
#dev.off()
```
The density graph produced are similar shaped to the Figure 6c in the paper but somehow the graphs in the paper are mixed up (the paper used Distance to T Cells as Distance to Tumor cells instead and vice versa)

```{r}
#===============================================================================
# Understand (Transform, Model, Visualize)
#===============================================================================
# generate some summary for the for the distance
# summary for tumor data distance as a whole
summary(tumor_data$Distance)
# summary for tumor data distance for H+
summary(tumor_data$Distance[tumor_data$Cell_type == "H+"])
# summary for tumor data distance for H-
summary(tumor_data$Distance[tumor_data$Cell_type == "H-"])

# summary for t cell data distance as a whole
summary(t_cell_data$Distance)
# summary for t cell data distance for H+
summary(t_cell_data$Distance[t_cell_data$Cell_type == "H+"])
# summary for t cell data distance for H-
summary(t_cell_data$Distance[t_cell_data$Cell_type == "H-"])

# generating some graph for interpretation
# tree map for cell types instead of pie charts because of this link
# https://www.data-to-viz.com/caveat/pie.html
#png(file = "./tree_map_cell_type.png")
table_cell_type <- data.frame(table(tumor_data$Cell_type))
ggplot(table_cell_type, aes(area = Freq, fill = Var1, 
                            label = paste(Var1, Freq, paste("(",format(round(Freq/7888*100, 2)), "%)", sep = ""), sep = "\n"))) +
  geom_treemap() +
  geom_treemap_text(color = "white",
                    place = "centre") +
  theme(legend.position = "none") +
  labs(title = "Tree map of cell types")
#dev.off

# tumor data
#plot(tumor_data$Cell_ID_Myeloid, tumor_data$Distance)
#plot(tumor_data$Cell_ID_Tumor, tumor_data$Distance)
#plot(tumor_data$Distance, tumor_data$Cell_ID_Myeloid) # scatterplot, too much points
#plot(tumor_data$Distance, tumor_data$Cell_ID_Tumor)
#plot(tumor_data$Cell_type, tumor_data$Distance)

# generate a boxplot with ggplot instead
# boxplot for distance to tumor cells
#png(file = "./boxplot_distance_to_tumor_cell.png")
ggplot(tumor_data, aes(x = fct_rev(Cell_type),
                       y = Distance,
                       fill = Cell_type)) +
  geom_boxplot() +
  xlab(" Cell type") +
  labs(title = "Boxplot distance to Tumor cells")
#dev.off()

# t cell data
#plot(t_cell_data$Cell_ID_Myeloid, t_cell_data$Distance)
#plot(t_cell_data$Cell_ID_T_Cell, t_cell_data$Distance)
#plot(t_cell_data$Distance, t_cell_data$Cell_ID_Myeloid) # scatterplot, too much points
#plot(t_cell_data$Distance, t_cell_data$Cell_ID_T_Cell)
#plot(t_cell_data$Cell_type, t_cell_data$Distance)

# boxplot for distance to t-cell
#png(file = "./boxplot_distance_to_t_cell.png")
ggplot(t_cell_data, aes(x = fct_rev(Cell_type),
                       y = Distance,
                       fill = Cell_type)) +
  geom_boxplot() +
  xlab(" Cell type") +
  labs(title = "Boxplot distance to T-cells")
#dev.off()

# ggplot of histogram and frequency line of individual instead of combine like in paper
# ggplot of histogram and frequency of tumor data
#png(file = "./histogram_density_line_distance_tumor_cell.png")
ggplot(tumor_data, aes(x = Distance)) +
  geom_freqpoly(color = "red") +
  geom_histogram(color = "black", alpha = 0.5) +
  ylab("Count") +
  labs(title = "Histogram + Density line of Distance to Tumor cell")
#dev.off()

# ggplot of histogram and density line of tumor data H+
#png(file = "./histogram_density_line_distance_tumor_cell_hpositive.png")
tumor_data_hpositive <- tumor_data[tumor_data$Cell_type == "H+",]
ggplot(tumor_data_hpositive, aes(x = Distance)) +
  geom_freqpoly(color = "#00BFC4") +
  geom_histogram(color = "#00BFC4", fill = "#00BFC4", alpha = 0.5) +
  ylab("Count") +
  labs(title = "Histogram + Density line of Distance to Tumor cell H+")
#dev.off()

# ggplot of histogram and density line of tumor data H-
#png(file = "./histogram_density_line_distance_tumor_cell_hnegative.png")
tumor_data_hnegative <- tumor_data[tumor_data$Cell_type == "H-",]
ggplot(tumor_data_hnegative, aes(x = Distance)) +
  geom_freqpoly(color = "#F8766D") +
  geom_histogram(color = "#F8766D", fill = "#F8766D", alpha = 0.5) +
  ylab("Count") +
  labs(title = "Histogram + Density line of Distance to Tumor cell H-")
#dev.off()

# ggplot of histogram and frequency of t cell data
#png(file = "./histogram_density_line_distance_t_cell.png")
ggplot(t_cell_data, aes(x = Distance)) +
  geom_freqpoly(color = "red") +
  geom_histogram(color = "black", alpha = 0.5) +
  ylab("Count") +
  labs(title = "Histogram + Density line of Distance to T-cell")
#dev.off()

# ggplot of histogram and density line of t cell data H+
#png(file = "./histogram_density_line_distance_t_cell_hpositive.png")
t_cell_data_hpositive <- t_cell_data[t_cell_data$Cell_type == "H+",]
ggplot(t_cell_data_hpositive, aes(x = Distance)) +
  geom_freqpoly(color = "#00BFC4") +
  geom_histogram(color = "#00BFC4", fill = "#00BFC4", alpha = 0.5) +
  ylab("Count") +
  labs(title = "Histogram + Density line of Distance to T cell H+")
#dev.off()

# ggplot of histogram and density line of t cell data H-
#png(file = "./histogram_density_line_distance_t_cell_hnegative.png")
t_cell_data_hnegative <- t_cell_data[t_cell_data$Cell_type == "H-",]
ggplot(t_cell_data_hnegative, aes(x = Distance)) +
  geom_freqpoly(color = "#F8766D") +
  geom_histogram(color = "#F8766D", fill = "#F8766D", alpha = 0.5) +
  ylab("Count") +
  labs(title = "Histogram + Density line of Distance to T cell H-")
#dev.off()
```

